### 2023-09-04 proofs/tex/subst-lemmas.tex
Grammar update: Γ = set

\begin{lemma}[Signature of a restricted substitution]
  \label{lemma:subst-restr-sig}
  If $[[Γ2 ⊢ σ : Γ1]]$ then $[[Γ2 ⊢ σ|varset : Ord {Γ1} ∩ varset]]$.
\end{lemma}
\begin{proof}
  Let us take an arbitrary $[[α± ∊ {Γ1} ∩ varset]]$.
  Since $[[α± ∊ {Γ1}]]$, $[[Γ2 ⊢ [σ]α± ]]$ by the signature of $[[σ]]$.

  Let us take an arbitrary $[[α± ∉ {Γ1} ∩ varset]]$.
  If $[[α± ∉ varset]]$ then $[[ [σ|varset]α± ]] = [[α±]]$ by definition
  of restriction. 
  If $[[α± ∊ varset \ {Γ1}]]$ then $[[ [σ|varset]α± ]] = [[ [σ]α± ]]$ by definition
  and $[[ [σ]α± ]] = [[α±]]$ by the signature of $[[σ]]$.
\end{proof}

@@

  If $[[α± ∊ varset]]$ then $[[α± ∉ {Γ1}]]$ by assumption.
  Then $[[ [σ|varset]α± = [σ]α±]]$ by definition of restricted substitution, and
  since $[[Γ2 ⊢ σ : Γ1]]$, we have $[[ [σ]α± = α± ]]$.
\end{proof}

\begin{corollary}[Application of a disjoint substitution]
  \label{corollary:subst-disj}
  Suppose that $[[σ]]$ is a substitution with signature $[[Γ2 ⊢ σ : Γ1]]$. Then
  \begin{itemize}
    \item [$+$] if $[[{Γ1} ∩ fv(iQ) = ∅]]$ then $[[ [σ] iQ = iQ]]$;
    \item [$-$] if $[[{Γ1} ∩ fv(iN) = ∅]]$ then $[[ [σ] iN = iN]]$.
  \end{itemize}
\end{corollary}

\begin{lemma}[Substitution range weakening]
  \label{lemma:subst-range-weakening}
  Suppose that $[[{Γ2} ⊆ {Γ'2}]]$ are contexts and 
  $[[σ]]$ is a substitution.
  Then $[[Γ2 ⊢ σ : Γ1]]$ implies $[[Γ'2 ⊢ σ : Γ1]]$.
\end{lemma}
\begin{proof}
  For any $[[α± ∊ {Γ1}]]$, 
  $[[Γ2 ⊢ σ : Γ1]]$ gives us  $[[ Γ2 ⊢ [σ]α± ]]$, which can be
  weakened to $[[ Γ'2 ⊢ [σ]α± ]]$ by \cref{lemma:wf-weakening}. 
  This way, $[[Γ'2 ⊢ σ : Γ1]]$.
\end{proof}

\begin{lemma}[]
  Suppose that $[[{Γ'} ⊆ {Γ}]]$,
  $[[σ1]]$ and $[[σ2]]$ are substitutions of signature $[[Γ ⊢ σi : Γ']]$.
  Then 
  \begin{enumerate}
    \item [$+$] for a type $[[Γ ⊢ iP]]$, if $[[Γ ⊢ [σ1]iP ≈ [σ2]iP]]$ then 
    $[[Γ ⊢ σ1 ≈ σ2 : fv iP ∩ {Γ'}]]$;
    \item [$-$] for a type $[[Γ ⊢ iN]]$, if $[[Γ ⊢ [σ1]iN ≈ [σ2]iN]]$ then
    $[[Γ ⊢ σ1 ≈ σ2 : fv iN ∩ {Γ'}]]$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Let us make an additional assumption that $[[σ1]]$, $[[σ2]]$, 
  and the mentioned types are normalized. If they are not,
  we normalize them first.

@@

  We prove it by induction on the structure of $[[iP]]$ and mutually, $[[iN]]$.
  Let us consider the shape of this type.
  \begin{caseof}
    \item $[[iP]] = [[α⁺]] \in [[Γ']]$.
      Then $[[Γ ⊢ σ1 ≈ σ2 : fv iP ∩ {Γ'}]]$ means $[[Γ ⊢ σ1 ≈ σ2 : {α⁺}]]$, 
      i.e. $[[Γ ⊢ [σ1]α⁺ ≈ [σ2]α⁺ ]]$, which holds by assumption.
    \item $[[iP]] = [[α⁺]] \in [[{Γ} \ {Γ'}]]$.
      Then $[[fv iP ∩ {Γ'}]] = [[∅]]$, 
      so $[[Γ ⊢ σ1 ≈ σ2 : fv iP ∩ {Γ'}]]$ holds vacuously.
    \item $[[iP]] = [[↓iN]]$.
      Then the induction hypothesis is applicable to type $[[iN]]$:
      \begin{enumerate}
        \item $[[iN]]$ is normalized,
        \item $[[Γ ⊢ iN]]$ by inversion of $[[Γ ⊢ ↓iN]]$,
        \item $[[Γ ⊢ [σ1]iN ≈ [σ2]iN]]$ holds by inversion of 
          $[[Γ ⊢ [σ1]↓iN ≈ [σ2]↓iN]]$, i.e. $[[Γ ⊢ ↓[σ1]iN ≈ ↓[σ2]iN]]$.
      \end{enumerate}
      This way, we obtain $[[Γ ⊢ σ1 ≈ σ2 : fv iN ∩ {Γ'}]]$, 
      which implies the required equivalence since 
      $[[fv iP ∩ {Γ'}]] = [[fv ↓iN ∩ {Γ'}]] = [[fv iN ∩ {Γ'}]]$.
    \item $[[iP]] = [[∃nas.iQ]]$
      Then the induction hypothesis is applicable to type $[[iQ]]$ 
      well-formed in context $[[Γ, nas]]$:
      \begin{enumerate}
        \item $[[{Γ'} ⊆ {Γ, nas}]]$ since $[[{Γ'} ⊆ {Γ}]]$,
        \item $[[Γ, nas ⊢ σi : Γ']]$ by weakening,
        \item $[[iQ]]$ is normalized,
        \item $[[Γ, nas ⊢ iQ]]$ by inversion of $[[Γ ⊢ ∃nas.iQ]]$,
        \item Notice that $[[ [σi]∃nas.iQ ]]$ is normalized, and thus, 
          $[[ [σ1]∃nas.iQ ≈ [σ2]∃nas.iQ]]$ implies 
          $[[ [σ1]∃nas.iQ = [σ2]∃nas.iQ ]]$
          (by \cref{lemma:subt-equiv-algorithmization}).).
          This equality means $[[ [σ1]iQ = [σ2]iQ ]]$, 
          which implies $[[Γ ⊢ [σ1]iQ ≈ [σ2]iQ]]$.
      \end{enumerate}
    \item $[[iN]] = [[iP → iM]]$
  \end{caseof}
\end{proof}

@@

\begin{lemma}[Substitution composition]
  \label{lemma:subst-composition}
    If $[[Γ'1 ⊢ σ1 : Γ1]]$, $[[Γ'2 ⊢ σ2 : Γ2]]$, 
    $[[{Γ1} ∩ {Γ'2} = ∅ ]]$ and $[[ {Γ1} ∩ {Γ2} = ∅ ]]$ then 
    $[[ σ2 ○ σ1 ]] = [[ (σ2 <=< σ1) ○ σ2 ]]$.
\end{lemma}

\begin{corollary}[Substitution composition commutativity]
  \label{corollary:subst-composition-commutativity}
  If $[[Γ'1 ⊢ σ1 : Γ1]]$, $[[Γ'2 ⊢ σ2 : Γ2]]$, and
  $[[ {Γ1} ∩ {Γ2} = ∅ ]]$, 
  $[[ {Γ1} ∩ {Γ'2} = ∅ ]]$, and
  $[[ {Γ'1} ∩ {Γ2} = ∅ ]]$ then 
  $[[ σ2 ○ σ1 ]] = [[ σ1 ○ σ2 ]]$.
\end{corollary}
\begin{proof}
  by \cref{lemma:subst-composition},
    $[[ σ2 ○ σ1 ]] = [[ (σ2 <=< σ1) ○ σ2 ]]$.
    Since the codomain of $[[σ1]]$ is $[[Γ'1]]$,
    and it is disjoint with the domain of $[[σ2]]$,
    $[[σ2 <=< σ1]] = [[σ1]]$.
\end{proof}

\begin{lemma}[Substitution domain weakening]
  \label{lemma:subst-domain-weakening}
  If $[[Γ2 ⊢ σ : Γ1]]$ then $[[Γ2, Γ' ⊢ σ : Γ1, Γ']]$
\end{lemma}
\begin{proof}
  If the variable $[[α±]]$ is in $[[Γ1]]$ then $[[Γ2 ⊢ [σ]α± ]]$ by assumption,
  and then $[[Γ2, Γ' ⊢ [σ]α± ]]$ by weakening.
  If the variable $[[α±]]$ is in $[[{Γ'} \ {Γ1}]]$ then $[[ [σ]α± = α± ]] \in [[Γ']] ⊆ [[Γ2, Γ']]$, 
  and thus, $[[Γ2, Γ' ⊢ α± ]]$.
\end{proof}

\begin{lemma}[Free variables after substitution]
  \label{lemma:subst-fv}
  Suppose that $[[Γ2 ⊢ σ : Γ1]]$, then
  \begin{enumerate}
    \item [$+$] for a type $[[iP]]$,
      the free variables of $[[ [σ]iP ]]$ are bounded in the following way:
     $[[fv(iP) \ {Γ1} ⊆ fv([σ]iP) ⊆ (fv(iP) \ {Γ1}) ∪ {Γ2}]]$;
    \item [$-$] for a type $[[iN]]$, 
      the free variables of $[[ [σ]iP ]]$ are bounded in the following way:
      $[[fv(iN) \ {Γ1} ⊆ fv([σ]iN) ⊆ (fv(iN) \ {Γ1}) ∪ {Γ2}]]$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  We prove it by structural induction on 
  $[[iP]]$ and mutually, on $[[iN]]$.
  \begin{caseof}
    \item $[[iP = α⁺]]$\\
      If $[[α⁺ ∊ {Γ1}]]$ then $[[Γ2 ⊢ [σ]α⁺]]$, and by 
      \cref{lemma:wf-soundness}, $[[fv([σ]α⁺) ⊆ {Γ2}]]$.
      $[[fv(α⁺) \ {Γ1}]] = [[∅]]$, so $[[fv([σ]iP) \ {Γ1}  ⊆ fv([σ]α⁺)]]$
      vacuously.

      If $[[α⁺ ∉ {Γ1}]]$ then $[[ [σ]α⁺ = α⁺ ]]$, and 
      $[[fv([σ]α⁺) = {α⁺}]] = [[{α⁺} \ {Γ1}]]$.

    \item $[[iP = ∃nas.iQ]]$\\
      Then we need to show that $[[fv([σ]iP) = fv([σ]iQ) \ {nas}]]$
      is a subset of $[[(fv(iP) \ {Γ1}) ∪ {Γ2}]]$ and a superset of
      $[[fv(iP) \ {Γ1}]]$.
      Notice that $[[fv(iP) = fv(iQ) \ {nas}]]$ by definition.
      This way, we need to show that 
      $[[fv(iQ) \ {nas} \ {Γ1} ⊆ fv([σ]iQ) \ {nas} ⊆ (fv(iQ) \ {nas} \ {Γ1}) ∪ {Γ2}]]$,

      By the induction hypothesis, 
      $[[fv([σ]iQ) ⊆ (fv(iQ) \ {Γ1}) ∪ {Γ2}]]$.
      So for the second inclusion, it suffices to show that
      $[[((fv(iQ) \ {Γ1}) ∪ {Γ2}) \ {nas} ⊆ (fv(iQ) \ {nas} \ {Γ1}) ∪ {Γ2}]]$,
      which holds by set theoretical reasoning.

      Also by the induction hypothesis,
      $[[fv(iQ) \ {Γ1} ⊆ fv([σ]iQ)]]$,
      and thus, by subtracting $[[{nas}]]$ from both sides,
      $[[fv(iQ) \ {nas} \ {Γ1} ⊆ fv([σ]iQ) \ {nas}]]$.

    \item The case $[[iN = ∀pas.iM]]$ is proved analogously.
    \item $[[iN = iP → iM]]$\\
      Then $[[fv([σ]iN) = fv([σ]iP) ∪ fv([σ]iM)]]$.
      By the induction hypothesis, 
      \begin{enumerate}
        \item $[[ fv(iP) \ {Γ1} ⊆ fv([σ]iP) ⊆ (fv(iP) \ {Γ1}) ∪ {Γ2}]]$ and
        \item $[[ fv(iM) \ {Γ1} ⊆ fv([σ]iM) ⊆ (fv(iM) \ {Γ1}) ∪ {Γ2}]]$.
      \end{enumerate}
      We unite these inclusions vertically and obtain
      $[[ fv(iP) \ {Γ1} ∪ fv(iM) \ {Γ1} ⊆ fv([σ]iN) ⊆ ((fv(iP) \ {Γ1}) ∪ {Γ2}) ∪ ((fv(iM) \ {Γ1}) ∪ {Γ2})]]$,
      which is equivalent to 
      $[[ (fv(iP) ∪ fv(iM)) \ {Γ1} ⊆ fv([σ]iN) ⊆ (fv(iP) ∪ fv(iM)) \ {Γ1} ∪ {Γ2}]]$.
      Since $[[fv(iP) ∪ fv(iM) = fv(iN)]]$, 
      $[[fv(iN) \ {Γ1} ⊆ fv([σ]iN) ⊆ (fv(iN) \ {Γ1}) ∪ {Γ2}]]$.
    \item The cases when $[[iP = ↓iM]]$ and $[[iN = ↑iQ]]$ are proved
      analogously
  \end{caseof}
\end{proof}

>>>>

\begin{lemma}[Signature of a restricted substitution]
  \label{lemma:subst-restr-sig}
  If $[[Γ2 ⊢ σ : Γ1]]$ then $[[Γ2 ⊢ σ|varset :  Γ1 ∩ varset]]$.
\end{lemma}
\begin{proof}
  Let us take an arbitrary $[[α± ∊ Γ1 ∩ varset]]$.
  Since $[[α± ∊ Γ1]]$, $[[Γ2 ⊢ [σ]α± ]]$ by the signature of $[[σ]]$.

  Let us take an arbitrary $[[α± ∉ Γ1 ∩ varset]]$.
  If $[[α± ∉ varset]]$ then $[[ [σ|varset]α± ]] = [[α±]]$ by definition
  of restriction. 
  If $[[α± ∊ varset \ Γ1]]$ then $[[ [σ|varset]α± ]] = [[ [σ]α± ]]$ by definition
  and $[[ [σ]α± ]] = [[α±]]$ by the signature of $[[σ]]$.
\end{proof}

@@

  If $[[α± ∊ varset]]$ then $[[α± ∉ Γ1]]$ by assumption.
  Then $[[ [σ|varset]α± = [σ]α±]]$ by definition of restricted substitution, and
  since $[[Γ2 ⊢ σ : Γ1]]$, we have $[[ [σ]α± = α± ]]$.
\end{proof}

\begin{corollary}[Application of a disjoint substitution]
  \label{corollary:subst-disj}
  Suppose that $[[σ]]$ is a substitution with signature $[[Γ2 ⊢ σ : Γ1]]$. Then
  \begin{itemize}
    \item [$+$] if $[[Γ1 ∩ fv(iQ) = ∅]]$ then $[[ [σ] iQ = iQ]]$;
    \item [$-$] if $[[Γ1 ∩ fv(iN) = ∅]]$ then $[[ [σ] iN = iN]]$.
  \end{itemize}
\end{corollary}

\begin{lemma}[Substitution range weakening]
  \label{lemma:subst-range-weakening}
  Suppose that $[[Γ2 ⊆ Γ'2]]$ are contexts and 
  $[[σ]]$ is a substitution.
  Then $[[Γ2 ⊢ σ : Γ1]]$ implies $[[Γ'2 ⊢ σ : Γ1]]$.
\end{lemma}
\begin{proof}
  For any $[[α± ∊ Γ1]]$, 
  $[[Γ2 ⊢ σ : Γ1]]$ gives us  $[[ Γ2 ⊢ [σ]α± ]]$, which can be
  weakened to $[[ Γ'2 ⊢ [σ]α± ]]$ by \cref{lemma:wf-weakening}. 
  This way, $[[Γ'2 ⊢ σ : Γ1]]$.
\end{proof}

\begin{lemma}[]
  Suppose that $[[Γ' ⊆ Γ]]$,
  $[[σ1]]$ and $[[σ2]]$ are substitutions of signature $[[Γ ⊢ σi : Γ']]$.
  Then 
  \begin{enumerate}
    \item [$+$] for a type $[[Γ ⊢ iP]]$, if $[[Γ ⊢ [σ1]iP ≈ [σ2]iP]]$ then 
    $[[Γ ⊢ σ1 ≈ σ2 : fv iP ∩ Γ']]$;
    \item [$-$] for a type $[[Γ ⊢ iN]]$, if $[[Γ ⊢ [σ1]iN ≈ [σ2]iN]]$ then
    $[[Γ ⊢ σ1 ≈ σ2 : fv iN ∩ Γ']]$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Let us make an additional assumption that $[[σ1]]$, $[[σ2]]$, 
  and the mentioned types are normalized. If they are not,
  we normalize them first.

@@

  We prove it by induction on the structure of $[[iP]]$ and mutually, $[[iN]]$.
  Let us consider the shape of this type.
  \begin{caseof}
    \item $[[iP]] = [[α⁺]] \in [[Γ']]$.
      Then $[[Γ ⊢ σ1 ≈ σ2 : fv iP ∩ Γ']]$ means $[[Γ ⊢ σ1 ≈ σ2 : {α⁺}]]$, 
      i.e. $[[Γ ⊢ [σ1]α⁺ ≈ [σ2]α⁺ ]]$, which holds by assumption.
    \item $[[iP]] = [[α⁺]] \in [[Γ \ Γ']]$.
      Then $[[fv iP ∩ Γ']] = [[∅]]$, 
      so $[[Γ ⊢ σ1 ≈ σ2 : fv iP ∩ Γ']]$ holds vacuously.
    \item $[[iP]] = [[↓iN]]$.
      Then the induction hypothesis is applicable to type $[[iN]]$:
      \begin{enumerate}
        \item $[[iN]]$ is normalized,
        \item $[[Γ ⊢ iN]]$ by inversion of $[[Γ ⊢ ↓iN]]$,
        \item $[[Γ ⊢ [σ1]iN ≈ [σ2]iN]]$ holds by inversion of 
          $[[Γ ⊢ [σ1]↓iN ≈ [σ2]↓iN]]$, i.e. $[[Γ ⊢ ↓[σ1]iN ≈ ↓[σ2]iN]]$.
      \end{enumerate}
      This way, we obtain $[[Γ ⊢ σ1 ≈ σ2 : fv iN ∩ Γ']]$, 
      which implies the required equivalence since 
      $[[fv iP ∩ Γ']] = [[fv ↓iN ∩ Γ']] = [[fv iN ∩ Γ']]$.
    \item $[[iP]] = [[∃nas.iQ]]$
      Then the induction hypothesis is applicable to type $[[iQ]]$ 
      well-formed in context $[[Γ, nas]]$:
      \begin{enumerate}
        \item $[[Γ' ⊆ Γ, nas]]$ since $[[Γ' ⊆ Γ]]$,
        \item $[[Γ, nas ⊢ σi : Γ']]$ by weakening,
        \item $[[iQ]]$ is normalized,
        \item $[[Γ, nas ⊢ iQ]]$ by inversion of $[[Γ ⊢ ∃nas.iQ]]$,
        \item Notice that $[[ [σi]∃nas.iQ ]]$ is normalized, and thus, 
          $[[ [σ1]∃nas.iQ ≈ [σ2]∃nas.iQ]]$ implies 
          $[[ [σ1]∃nas.iQ = [σ2]∃nas.iQ ]]$
          (by \cref{lemma:subt-equiv-algorithmization}).).
          This equality means $[[ [σ1]iQ = [σ2]iQ ]]$, 
          which implies $[[Γ ⊢ [σ1]iQ ≈ [σ2]iQ]]$.
      \end{enumerate}
    \item $[[iN]] = [[iP → iM]]$
  \end{caseof}
\end{proof}

@@

\begin{lemma}[Substitution composition]
  \label{lemma:subst-composition}
    If $[[Γ'1 ⊢ σ1 : Γ1]]$, $[[Γ'2 ⊢ σ2 : Γ2]]$, 
    $[[Γ1 ∩ Γ'2 = ∅ ]]$ and $[[ Γ1 ∩ Γ2 = ∅ ]]$ then 
    $[[ σ2 ○ σ1 ]] = [[ (σ2 <=< σ1) ○ σ2 ]]$.
\end{lemma}

\begin{corollary}[Substitution composition commutativity]
  \label{corollary:subst-composition-commutativity}
  If $[[Γ'1 ⊢ σ1 : Γ1]]$, $[[Γ'2 ⊢ σ2 : Γ2]]$, and
  $[[ Γ1 ∩ Γ2 = ∅ ]]$, 
  $[[ Γ1 ∩ Γ'2 = ∅ ]]$, and
  $[[ Γ'1 ∩ Γ2 = ∅ ]]$ then 
  $[[ σ2 ○ σ1 ]] = [[ σ1 ○ σ2 ]]$.
\end{corollary}
\begin{proof}
  by \cref{lemma:subst-composition},
    $[[ σ2 ○ σ1 ]] = [[ (σ2 <=< σ1) ○ σ2 ]]$.
    Since the codomain of $[[σ1]]$ is $[[Γ'1]]$,
    and it is disjoint with the domain of $[[σ2]]$,
    $[[σ2 <=< σ1]] = [[σ1]]$.
\end{proof}

\begin{lemma}[Substitution domain weakening]
  \label{lemma:subst-domain-weakening}
  If $[[Γ2 ⊢ σ : Γ1]]$ then $[[Γ2, Γ' ⊢ σ : Γ1, Γ']]$
\end{lemma}
\begin{proof}
  If the variable $[[α±]]$ is in $[[Γ1]]$ then $[[Γ2 ⊢ [σ]α± ]]$ by assumption,
  and then $[[Γ2, Γ' ⊢ [σ]α± ]]$ by weakening.
  If the variable $[[α±]]$ is in $[[Γ' \ Γ1]]$ then $[[ [σ]α± = α± ]] \in [[Γ']] ⊆ [[Γ2, Γ']]$, 
  and thus, $[[Γ2, Γ' ⊢ α± ]]$.
\end{proof}

\begin{lemma}[Free variables after substitution]
  \label{lemma:subst-fv}
  Suppose that $[[Γ2 ⊢ σ : Γ1]]$, then
  \begin{enumerate}
    \item [$+$] for a type $[[iP]]$,
      the free variables of $[[ [σ]iP ]]$ are bounded in the following way:
     $[[fv(iP) \ Γ1 ⊆ fv([σ]iP) ⊆ (fv(iP) \ Γ1) ∪ Γ2]]$;
    \item [$-$] for a type $[[iN]]$, 
      the free variables of $[[ [σ]iP ]]$ are bounded in the following way:
      $[[fv(iN) \ Γ1 ⊆ fv([σ]iN) ⊆ (fv(iN) \ Γ1) ∪ Γ2]]$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  We prove it by structural induction on 
  $[[iP]]$ and mutually, on $[[iN]]$.
  \begin{caseof}
    \item $[[iP = α⁺]]$\\
      If $[[α⁺ ∊ Γ1]]$ then $[[Γ2 ⊢ [σ]α⁺]]$, and by 
      \cref{lemma:wf-soundness}, $[[fv([σ]α⁺) ⊆ Γ2]]$.
      $[[fv(α⁺) \ Γ1]] = [[∅]]$, so $[[fv([σ]iP) \ Γ1  ⊆ fv([σ]α⁺)]]$
      vacuously.

      If $[[α⁺ ∉ Γ1]]$ then $[[ [σ]α⁺ = α⁺ ]]$, and 
      $[[fv([σ]α⁺) = {α⁺}]] = [[{α⁺} \ Γ1]]$.

    \item $[[iP = ∃nas.iQ]]$\\
      Then we need to show that $[[fv([σ]iP) = fv([σ]iQ) \ {nas}]]$
      is a subset of $[[(fv(iP) \ Γ1) ∪ Γ2]]$ and a superset of
      $[[fv(iP) \ Γ1]]$.
      Notice that $[[fv(iP) = fv(iQ) \ {nas}]]$ by definition.
      This way, we need to show that 
      $[[fv(iQ) \ {nas} \ Γ1 ⊆ fv([σ]iQ) \ {nas} ⊆ (fv(iQ) \ {nas} \ Γ1) ∪ Γ2]]$,

      By the induction hypothesis, 
      $[[fv([σ]iQ) ⊆ (fv(iQ) \ Γ1) ∪ Γ2]]$.
      So for the second inclusion, it suffices to show that
      $[[((fv(iQ) \ Γ1) ∪ Γ2) \ {nas} ⊆ (fv(iQ) \ {nas} \ Γ1) ∪ Γ2]]$,
      which holds by set theoretical reasoning.

      Also by the induction hypothesis,
      $[[fv(iQ) \ Γ1 ⊆ fv([σ]iQ)]]$,
      and thus, by subtracting $[[{nas}]]$ from both sides,
      $[[fv(iQ) \ {nas} \ Γ1 ⊆ fv([σ]iQ) \ {nas}]]$.

    \item The case $[[iN = ∀pas.iM]]$ is proved analogously.
    \item $[[iN = iP → iM]]$\\
      Then $[[fv([σ]iN) = fv([σ]iP) ∪ fv([σ]iM)]]$.
      By the induction hypothesis, 
      \begin{enumerate}
        \item $[[ fv(iP) \ Γ1 ⊆ fv([σ]iP) ⊆ (fv(iP) \ Γ1) ∪ Γ2]]$ and
        \item $[[ fv(iM) \ Γ1 ⊆ fv([σ]iM) ⊆ (fv(iM) \ Γ1) ∪ Γ2]]$.
      \end{enumerate}
      We unite these inclusions vertically and obtain
      $[[ fv(iP) \ Γ1 ∪ fv(iM) \ Γ1 ⊆ fv([σ]iN) ⊆ ((fv(iP) \ Γ1) ∪ Γ2) ∪ ((fv(iM) \ Γ1) ∪ Γ2)]]$,
      which is equivalent to 
      $[[ (fv(iP) ∪ fv(iM)) \ Γ1 ⊆ fv([σ]iN) ⊆ (fv(iP) ∪ fv(iM)) \ Γ1 ∪ Γ2]]$.
      Since $[[fv(iP) ∪ fv(iM) = fv(iN)]]$, 
      $[[fv(iN) \ Γ1 ⊆ fv([σ]iN) ⊆ (fv(iN) \ Γ1) ∪ Γ2]]$.
    \item The cases when $[[iP = ↓iM]]$ and $[[iN = ↑iQ]]$ are proved
      analogously
  \end{caseof}
\end{proof}

### 2023-09-04 proofs/tex/upgrade-lemmas.tex
Grammar update: Γ = set

Let us consider a type $[[iP]]$ well-formed in $[[Γ]]$.
Some of its $[[Γ]]$-supertypes are also well-formed in a smaller context $[[{Δ} ⊆ {Γ}]]$.
The upgrade is the operation that returns the least of such supertypes.

@@

    $ 
    \begin{aligned}
        [[fv iQ]] &\subseteq [[fv [pnbs/pnas]iP ∩ fv [pncs/pnas]iP]]
                  &&\text{Since by \cref{lemma:fv-propagation}, 
                         $[[fv iQ ⊆ fv [pnbs/pnas]iP]]$ and
                         $[[fv iQ ⊆ fv [pncs/pnas]iP]]$}\\
                  &\subseteq [[ ((fv iP \ {pnas}) ∪ {pnbs}) ∩ ((fv iP \ {pnas}) ∪ {pncs})]]\\
                  &= [[ (fv iP \ {pnas}) ∩ (fv iP \ {pnas}) ]]
                  &&\text{since $[[pnbs]]$ and $[[pncs]]$ are fresh}\\
                  &= [[ fv iP \ {pnas} ]]\\
                  &\subseteq [[ {Γ} \ {pnas} ]]
                  &&\text{since $[[iP]]$ is well-formed in $[[Γ]]$}\\
                  &\subseteq [[ {Δ} ]]\\
    \end{aligned}
    $\\
    This way, by \cref{lemma:wf-ctxt-equiv}, $[[Δ ⊢ iQ]]$.

>>>>

Let us consider a type $[[iP]]$ well-formed in $[[Γ]]$.
Some of its $[[Γ]]$-supertypes are also well-formed in a smaller context $[[{Δ} ⊆ Γ]]$.
The upgrade is the operation that returns the least of such supertypes.

@@

    $ 
    \begin{aligned}
        [[fv iQ]] &\subseteq [[fv [pnbs/pnas]iP ∩ fv [pncs/pnas]iP]]
                  &&\text{Since by \cref{lemma:fv-propagation}, 
                         $[[fv iQ ⊆ fv [pnbs/pnas]iP]]$ and
                         $[[fv iQ ⊆ fv [pncs/pnas]iP]]$}\\
                  &\subseteq [[ ((fv iP \ {pnas}) ∪ {pnbs}) ∩ ((fv iP \ {pnas}) ∪ {pncs})]]\\
                  &= [[ (fv iP \ {pnas}) ∩ (fv iP \ {pnas}) ]]
                  &&\text{since $[[pnbs]]$ and $[[pncs]]$ are fresh}\\
                  &= [[ fv iP \ {pnas} ]]\\
                  &\subseteq [[ Γ \ {pnas} ]]
                  &&\text{since $[[iP]]$ is well-formed in $[[Γ]]$}\\
                  &\subseteq [[ {Δ} ]]\\
    \end{aligned}
    $\\
    This way, by \cref{lemma:wf-ctxt-equiv}, $[[Δ ⊢ iQ]]$.
